### 2023-02-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@c9e3dd5b428ea22dc5ecfea7425b6f717e3897de 🚀

Cataloging species interactions across space is a gargantuan task. At
the core of this challenge lies the spatiotemporal variability of
ecological networks (\textbf{Poisot2012DisSpe?};
\textbf{Poisot2015SpeWhy?}), which makes documenting the location and
timing of interactions difficult. Indeed, it is not sufficient to know
that two species have the biological capacity to interact to infer the
realization of their interaction at a specific time and space (Dunne
2006). Taking food webs as an example, a predator species and its
potential prey must first co-occur on the same territory in order for a
trophic interaction to take place (\textbf{Blanchet2020CooNot?}). They
must then encounter, which is conditional on their relative abundances
in the ecosystem and the matching of their phenology
(\textbf{Poisot2015SpeWhy?}). Finally, the interaction occurs only if
the predators have a desire to consume their prey and are able to
capture and ingest them (\textbf{Pulliam1974TheOpt?}). Environmental
(e.g.~temperature and presence of shelters) and biological
(e.g.~physiological state of both species and availability of other prey
species) factors contribute to this variability by impacting species
co-occurrence (Araujo \emph{et al.} 2011) and the realization of their
interactions (\textbf{Poisot2015SpeWhy?}). In this context, it is
unsurprising that computational methods are being developed in ecology
to help alleviate the colossal sampling efforts required to document
species interactions across time and space
(\textbf{Strydom2021RoaPre?}). Having a better portrait of species
interactions and the emerging structure of their food webs is important
since it lays the groundwork for understanding the functioning,
dynamics, and resilience of ecosystems worldwide (e.g., Proulx \emph{et
al.} 2005; Pascual \emph{et al.} 2006; Delmas \emph{et al.} 2019).

The recognition of the intrinsic variability of species interactions and
the emergence of numerical methods have led ecologists to rethink their
representation of ecological networks, slowly moving from a binary to a
probabilistic view of species interactions (Poisot \emph{et al.} 2016).
This has several benefits. For example, probabilities represent the
limit of our knowledge about species interactions and can inform us
about the expected number of interactions and emerging network
properties despite this limited knowledge (Poisot \emph{et al.} 2016).
They are also very helpful in predicting the spatial distribution of
species within networks (\textbf{Cazelles2016TheSpe?}) and the temporal
variability of interactions (Poisot \emph{et al.} 2015), generating new
ecological data (e.g., Strydom \emph{et al.} 2022), and identifying
priority sampling locations of species interactions (see
\textbf{Andrade-Pacheco2020FinHot?} for an ecological example of a
sampling optimization problem). Moreover, the high rate of false
negatives in ecological network data, resulting from the difficulty of
witnessing interactions between rare species, makes it hard to interpret
non-observations of species interactions ecologically
(\textbf{Catchen2023MisLin?}). Using probabilities instead of yes-no
interactions accounts for these observation errors; in that case, only
forbidden interactions (\textbf{Jordano2003InvPro?};
\textbf{Olesen2010MisFor?}) would have a probability value of zero (but
see \textbf{Gonzalez-Varo2016LabLim?}). Many measures have been
developed to describe the structure (Poisot \emph{et al.} 2016) and
diversity (Ohlmann \emph{et al.} 2019; Godsoe \emph{et al.} n.d.) of
probabilistic interactions, which shows the potential of this framework
in the study of a variety of ecological phenomena.

@@

\textbf{Papers:} (\textbf{Strydom2022FooWeb?})

@@

\textbf{Papers:} there might be something in these McLeod \emph{et al.}
(2020); (\textbf{McLeod2021SamAsy?}); (\textbf{Wood2015EffSpa?})

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Araujo2011UsiSpe}{}%
Araujo, M.B., Rozenfeld, A., Rahbek, C. \& Marquet, P.A. (2011). Using
species co-occurrence networks to assess the impacts of climate change.
\emph{Ecography}, 34, 897--908.

@@

@@

@@

\leavevmode\hypertarget{ref-GodsoeSpeInt}{}%
Godsoe, W., Murray, R. \& Iritani, R. (n.d.). Species interactions and
diversity: A unified framework using Hill numbers. \emph{Oikos}, n/a,
e09282.

@@

@@

@@

@@

@@

>>>>

Cataloging species interactions across space is a gargantuan task. At
the core of this challenge lies the spatiotemporal variability of
ecological networks (Poisot \emph{et al.} 2012, 2015), which makes
documenting the location and timing of interactions difficult. Indeed,
it is not sufficient to know that two species have the biological
capacity to interact to infer the realization of their interaction at a
specific time and space (Dunne 2006). Taking food webs as an example, a
predator species and its potential prey must first co-occur on the same
territory in order for a trophic interaction to take place (Blanchet
\emph{et al.} 2020). They must then encounter, which is conditional on
their relative abundances in the ecosystem and the matching of their
phenology (Poisot \emph{et al.} 2015). Finally, the interaction occurs
only if the predators have a desire to consume their prey and are able
to capture and ingest them (Pulliam 1974). Environmental
(e.g.~temperature and presence of shelters) and biological
(e.g.~physiological state of both species and availability of other prey
species) factors contribute to this variability by impacting species
co-occurrence (Araujo \emph{et al.} 2011) and the realization of their
interactions (Poisot \emph{et al.} 2015). In this context, it is
unsurprising that computational methods are being developed in ecology
to help alleviate the colossal sampling efforts required to document
species interactions across time and space (Strydom \emph{et al.} 2021).
Having a better portrait of species interactions and the emerging
structure of their food webs is important since it lays the groundwork
for understanding the functioning, dynamics, and resilience of
ecosystems worldwide (e.g., Proulx \emph{et al.} 2005; Pascual \emph{et
al.} 2006; Delmas \emph{et al.} 2019).

The recognition of the intrinsic variability of species interactions and
the emergence of numerical methods have led ecologists to rethink their
representation of ecological networks, slowly moving from a binary to a
probabilistic view of species interactions (Poisot \emph{et al.} 2016).
This has several benefits. For example, probabilities represent the
limit of our knowledge about species interactions and can inform us
about the expected number of interactions and emerging network
properties despite this limited knowledge (Poisot \emph{et al.} 2016).
They are also very helpful in predicting the spatial distribution of
species within networks (Cazelles \emph{et al.} 2016) and the temporal
variability of interactions (Poisot \emph{et al.} 2015), generating new
ecological data (e.g., Strydom \emph{et al.} 2022), and identifying
priority sampling locations of species interactions (see Andrade-Pacheco
\emph{et al.} 2020 for an ecological example of a sampling optimization
problem). Moreover, the high rate of false negatives in ecological
network data, resulting from the difficulty of witnessing interactions
between rare species, makes it hard to interpret non-observations of
species interactions ecologically (Catchen \emph{et al.} 2023). Using
probabilities instead of yes-no interactions accounts for these
observation errors; in that case, only forbidden interactions (Jordano
\emph{et al.} 2003; Olesen \emph{et al.} 2010) would have a probability
value of zero (but see Gonzalez-Varo \& Traveset 2016). Many measures
have been developed to describe the structure (Poisot \emph{et al.}
2016) and diversity (Ohlmann \emph{et al.} 2019; \textbf{GodsoeSpeInt?})
of probabilistic interactions, which shows the potential of this
framework in the study of a variety of ecological phenomena.

@@

\textbf{Papers:} Strydom \emph{et al.} (2022)

@@

\textbf{Papers:} there might be something in these McLeod \emph{et al.}
(2020); McLeod \emph{et al.} (2021); Wood \emph{et al.} (2015)

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Andrade-Pacheco2020Finding}{}%
Andrade-Pacheco, R., Rerolle, F., Lemoine, J., Hernandez, L., Meïté, A.,
Juziwelo, L., \emph{et al.} (2020). Finding hotspots: Development of an
adaptive spatial sampling approach. \emph{Scientific Reports}, 10,
10939.

\leavevmode\hypertarget{ref-Araujo2011UsiSpe}{}%
Araujo, M.B., Rozenfeld, A., Rahbek, C. \& Marquet, P.A. (2011). Using
species co-occurrence networks to assess the impacts of climate change.
\emph{Ecography}, 34, 897--908.

@@

\leavevmode\hypertarget{ref-Blanchet2020Cooccurrencea}{}%
Blanchet, F.G., Cazelles, K. \& Gravel, D. (2020). Co-occurrence is not
evidence of ecological interactions. \emph{Ecology Letters}, 23,
1050--1063.

@@

\leavevmode\hypertarget{ref-Catchen2023Missinga}{}%
Catchen, M.D., Poisot, T., Pollock, L.J. \& Gonzalez, A. (2023). The
missing link: Discerning true from false negatives when sampling species
interaction networks.

\leavevmode\hypertarget{ref-Cazelles2016Theorya}{}%
Cazelles, K., Araujo, M.B., Mouquet, N. \& Gravel, D. (2016). A theory
for species co-occurrence in interaction networks. \emph{Theoretical
Ecology}, 9, 39--48.

@@

\leavevmode\hypertarget{ref-Gonzalez-Varo2016Labilea}{}%
Gonzalez-Varo, J.P. \& Traveset, A. (2016). The Labile Limits of
Forbidden Interactions. \emph{Trends in Ecology \& Evolution}, 31,
700--710.

@@

\leavevmode\hypertarget{ref-Jordano2003Invarianta}{}%
Jordano, P., Bascompte, J. \& Olesen, J.M. (2003). Invariant properties
in coevolutionary networks of plantanimal interactions. \emph{Ecology
Letters}, 6, 69--81.

\leavevmode\hypertarget{ref-McLeod2021Sampling}{}%
McLeod, A., Leroux, S.J., Gravel, D., Chu, C., Cirtwill, A.R., Fortin,
M.-J., \emph{et al.} (2021). Sampling and asymptotic network properties
of spatial multi-trophic networks. \emph{Oikos}, n/a.

@@

\leavevmode\hypertarget{ref-Olesen2010Missing}{}%
Olesen, J.M., Bascompte, J., Dupont, Y.L., Elberling, H., Rasmussen, C.
\& Jordano, P. (2010). Missing and forbidden links in mutualistic
networks. \emph{Proceedings of the Royal Society B: Biological
Sciences}, 278, 725--732.

@@

\leavevmode\hypertarget{ref-Poisot2012Dissimilaritya}{}%
Poisot, T., Canard, E., Mouillot, D., Mouquet, N. \& Gravel, D. (2012).
The dissimilarity of species interaction networks. \emph{Ecology
Letters}, 15, 1353--1361.

@@

\leavevmode\hypertarget{ref-Pulliam1974Theory}{}%
Pulliam, H.R. (1974). On the Theory of Optimal Diets. \emph{The American
Naturalist}, 108, 59--74.

@@

\leavevmode\hypertarget{ref-Wood2015Effects}{}%
Wood, S.A., Russell, R., Hanson, D., Williams, R.J. \& Dunne, J.A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, 5, 3769--3782.
